The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Finished results chapter and summary.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@634 4525493e-7705-40b1-a816-d608a930855b

\section{Results\label{Results}}
DAS was deployed to work with the following CMS data-services:
\begin{itemize}
\item DBS \cite{DBS}, Data Bookkeeping System, which collect information
about CMS meta-data;
\item PhEDEx \cite{PhEDEx}, Physics Experiment Data Export project, which
provides the data placement and the file transfer system for the CMS experiment;
\item SiteDB \cite{SiteDB} which 
records the CMS resources at the site, the resources pledged for the 
future, and it keeps track of CMS personnel at each site, including the 
roles and duties they fulfill in the collaboration;
\item RunSummary DB \cite{RunSummary} which collects information about run and triggers
conditions during data-taking;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard} which monitors data collected from the 
distributed computing systems of the CMS experiment and LHC virtual organization
in general.
\end{itemize}
Each data-service has its own scope, size and lifetime. For instance the data
in PhEDEx are transient due to constant migration of CMS data, 
the DBS system were divided into dozen of individual instances, 
who has been used by different production teams for various bookkeeping tasks.
Figure \ref{db_size} shows current size of meta-data from aforementioned data-services.
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{db_size.pdf}
\caption{
Current size of meta-data for CMS data-services participated in DAS, before
official data-taking. We anticipate to collect $\sim$500GB of meta-data
per year once LHC starts operating.
}
\label{db_size}
\end{figure}
%Typical
%size of individual DBS instance was around of fwe GBs, with overall DBS size of
%around of 50 GBs.\footnote{This include table and index sizes.}

At the moment we don't have exact numbers for amount of data which will stored to DAS,
but based on test performed during 
Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}
exercise we produced 0.3PB of data which
translate into $\sim50$ GB of meta-data stored in DBS system. Therefore
we can anticipate to collect $\sim500$GB of meta-data per year during data-taking period
at LHC.
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have signicant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

Preliminary results of DAS performance are shown in Table \ref{DAS_benchmark}.
We benchmark time required to collect and aggregate block object information
from DBS and PhEDEx CMS data-services.

@@

Elapsed time measured during this test consist of the following componenets:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here the {\it retrieval time} is a time required to access data from remote data-service,
{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
is a time to convert notations used by data-service to DAS ones for every object
we parse, {\it cache insertion} and {\it indexing time} represents time spend to add objects into
the DAS cache, {\it aggregation time} is required to merge objects into DAS records based
on their common key (block name in this case) and {\it output creation time}
is a time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, were only applied to
final DAS step and not to time spent in individual data-services.

@@

The individual DAS components (DBS and PhEDEx) shown roughly the same performance.
Although the time of PhEDEx component was divided among: retrieval + parsing 
time $\sim$110 sec and merging step $\sim$50 sec.

As you can see from Table \ref{DAS_benchmark},
the time spent to read DAS records (last column) from the cache was quite
reasonable, roughly 1 second for both systems. While final time to
get DAS records on disk was about 45 seconds. This was mostly due to I/O and
conversion between binary data format used in database (BSON) to JSON.
All operations were performed on a single CPU core and obviously can be done
in parallel.

Based on this results we measured that we can achieve 7000 docs/sec rate
for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for reading operation
in DAS, which is suitable for our 
needs.\footnote{We were satisfied with performance of the DAS prototype,
even though it throughput can be increased by few times
using C++ DB driver instead of python one used in this system.}

\section{Summary}
We presented new data aggregation service (DAS) developed for CMS High-Energy experiment
at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
aggregation layer on top of the existing relational and non-relation data-services
mostly in real time fashion. All the data were retrieved on demand basis,
while data pre-fetching of most common queries is in our plans. We developed
prototype in python for DAS system which is under commissioning phase right now 
and performed performance studies shown in this paper. 
The CMS experiment is started to
collect data from November 2009 and we expect to write about PB of data each
year to tape. In addition, the Monte-Carlo samples will be produced at this scale.
Based on test studies performed in CMS, we expect that total size of
meta-data produced by experiment each year will be of the order of
$\sim500$GB. The DAS should sustain such load and will provide generic 
``data-discovery'' service for CMS experiment in years to come.

@@

>>>>

\section{Results\label{Results}}
In this section we discuss preliminary restults achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS file transfer system and data-service;
\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiar properties. 
For instance, the data in PhEDEx are transient due to constant 
migration of CMS data; the DBS system were divided into dozen 
of individual instances with own size and lifetime, who has been used by different production 
teams for various bookkeeping tasks; the RunSummary data-service
operates on private network accessible at CMS online operation center.
Figure \ref{db_size} shows the current size of meta-data from aforementioned 
data-services.
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{db_size.pdf}
\caption{
Current size of meta-data for CMS data-services participated in DAS, before
official data-taking. 
%We anticipate to collect $\sim$500GB of meta-data
%per year during normal LHC operations.
}
\label{db_size}
\end{figure}

At the moment we don't have exact numbers for amount of data which will stored 
in those data-service and therefor accessible in DAS. Although we project to
accumulate of the order of 500 GB of meta-data per year during normal LHC and CMS
operations.\footnote{This number is based on tests 
performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have signicant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

To test DAS performance we measured elapsed time and CPU/RAM utilization
required to collect and aggregate block object information from DBS and 
PhEDEx CMS data-services. Preliminary results of those measurements can
be found in Table \ref{DAS_benchmark}.

@@

Please note that initial DAS implementation was done in Python \cite{Python}
and all tests were performed on x64 Linux platform with 16GB of RAM where
1 CPU was used by DAS server and 1 CPU by MongoDB database server. The CPU utilization
of DAS server was 67\% in average and RAM usage was below 35MB for
all tests. All tests were performed multiple times to avoid fluctuations. 
The elapsed time measured during those tests consists of the 
following componenets:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here the {\it retrieval time} is a time required to access data from remote data-service,
{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
is a time required to convert notations used by data-service to DAS ones for every object
we parse, {\it cache insertion} and {\it indexing time} represents time spend to add objects into
the DAS cache, {\it aggregation time} is required to merge objects into DAS records based
on their common key (block name in this case) and {\it output creation time}
is a time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, were only applied to
the one of the sub-system (either DBS or PhEDEx) since the first one was injected into
fresh database, while the second required the merging step.

@@

The individual tests of DAS components (DBS and PhEDEx) shown roughly 
the same performance. Although the time of PhEDEx component was bigger 
due to longer retrieval and parsing time.
The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
among: retrieval + parsing time $\sim$110 sec and merging step $\sim$50 sec.

As you can see from Table \ref{DAS_benchmark},
the time spent to access individual components (last column) was quite
reasonable, roughly 1 second for both systems. While final time to
get DAS records on disk was about 45 seconds. This was mostly due to I/O and
conversion between binary data format used by MongoDB, BSON, to DAS
data format, JSON.

Based on this results we measured that we can achieve 7000 docs/sec rate
for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for reading operation
in DAS, which is suitable for our needs. Since all operations were performed 
on a single CPU core a further improvements are expected. Also, a necesarry
tuning can be done by using native C++ MongoDB driver instead of python one.

\section{Summary}
We presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
aggregation layer on top of the existing relational and non-relation data-services
mostly in real time fashion. All of the data were retrieved on demand basis,
while data pre-fetching of most common queries is in our plans. We developed
prototype in python for DAS system which is under commissioning phase right now 
and performed performance studies shown in this paper. 
The CMS experiment is started to collect its data in December of 2009 and 
we expect to accumulate a few PB of real data each year to tape. 
In addition, the Monte-Carlo samples will be produced at this scale.
Based on test studies performed in CMS, we expect that total size of
meta-data produced by experiment each year will be of the order of
$\sim500$GB. The DAS should sustain such load and will provide generic 
``data-discovery'' service for CMS experiment in years to come.

@@

\bibitem{Python} Python programming language, http://www.python.org
